---
title: "6PM_assign_6"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## The power curve
Let $z = \sqrt{n}\bar{X}$, then under $H_0$, $z \sim N(0,1)$, 

under $H_1$, $\sqrt{n}(\bar{X} - \mu) \sim N(0,1)$.

$$\begin{align}
power &= Pr(reject \ H_0 |\ H_1 \ \text{is true}) \\
&= Pr(z > 1.96 |\ H_1) + Pr(z < -1.96 |\ H_1) \\
&= 1 - Pr(\sqrt{n}\bar{X} < 1.96 |\ H_1) + Pr(\sqrt{n}\bar{X} < -1.96 |\ H_1)) \\
&= 1 - Pr(\sqrt{n}(\bar{X} - \mu)> 1.96- \sqrt{n}\mu|\ H_1) \\
&+ Pr(\sqrt{n}(\bar{X} - \mu) < -1.96 - \sqrt{n}\mu|\ H_1) \\
&= 1 - \Phi(1.96 - \sqrt{n}\mu) + \Phi(-1.96 - \sqrt{n}\mu)
\end{align}$$
where $\Phi(\cdot)$ is the cdf of standard normal.

```{r , results = "hide", message=FALSE}
library(tidyverse) #everything
library(lmtest) # coeftest
library(sandwich) # vcovHC
library(dplyr)
library(broom) # tidy the output
library(knitr) # For knitting document and include_graphics function
library(png)      # For grabbing the dimensions of png files
library(magrittr)
library(png) 
library(grid) 
library(gridExtra)

rm(list = ls())      # clean the workspace

S <- 100
size_gr <- expand.grid(1:S, n=(2:30)^2, mu=seq(-5,5,by = 0.25))
alpha = 0.05
test <- function(n,mu){
  X = rnorm(n,mu,1)
  z = sqrt(n)*mean(X)
  test <- data.frame(
    n = n,
    mu = mu,
    reject = (abs(z)> qnorm(1 - alpha/2))
  )
  return(test)
}

simu_gr <- mapply(test, n=size_gr$n, mu = size_gr$mu, SIMPLIFY = FALSE) %>%
  bind_rows() %>%
  as_data_frame()

reject <- simu_gr %>%
  group_by(n,mu) %>%
  summarise(
  ave_prob = mean(reject)
)
```

## Heatmap of Average Rejection Rate

```{r, heatmap}
## Heatmap
ggplot(reject, aes(x=mu, y = factor(n))) +
  geom_tile(aes(fill=ave_prob))
```

## Comments

- Under $H_0$, the probability of rejecting is just the size, which equals 0.05.
- Under $H_1$, as the sample size increases or the deviation from the null increases, power increases.

## Power Curve under 4 n's

```{r, n}
theo_pow <- function(n, mu){
  y = 1 - pnorm(1.96 - sqrt(n)*mu, mean=0, sd=1) + pnorm(-1.96 - sqrt(n)*mu, mean=0, sd=1)
  return(y)
}
power_n <- simu_gr %>%
  filter(n == c("36","81","324","625")) %>%
  group_by(n,mu) %>%
  summarise(
    ave_prob = mean(reject)
  )%>%
  mutate(power = map2_dbl(n,mu, ~theo_pow(.x,.y)))

ggplot(power_n, aes(x = mu, colour = factor(n))) +
  geom_line(aes(y = ave_prob)) + 
  geom_line(aes(y=power), linetype = 3)
```

## Comments

- As the sample size increases, the power curve shrinks, and the power curve gets closer to the theoretical one.
- After the sample size gets large enough (say 400), there's little improvement in power afterwards.

## Power Curve under 4 mu's

```{r,mu}
power_mu <- simu_gr %>%
  filter(mu== c("0","0.25","0.5","2")) %>%
  group_by(n,mu) %>%
  summarise(
    ave_prob = mean(reject)
  ) %>%
  mutate(power = map2_dbl(n,mu, ~theo_pow(.x,.y)))

ggplot(power_mu, aes(x = n, colour = factor(mu))) +
  geom_line(aes(y = ave_prob)) + 
  geom_line(aes(y=power), linetype = 3)
```

## Comments

- Under $H_0$, the rejection rate in theory should be 0.05 regardless of sample size, and simulated result is consistent with this fact.
- Under $H_1$, as the samples size increases, the power converges to 1. Also the larger the $\mu$ is, the faster the power converges.

```{r , results = "hide", message=FALSE}
n_pow = (2:30)^2
pows_df <- data_frame(Type = c("Turtle","Good student","Rabbit"),
                      power = c(0.4,0.5,0.6)) %>%
  mutate(mu = map(power, ~data_frame(n = n_pow, mu = 2/n_pow^.))) %>%
  unnest(mu)

sim_fun <- function(n,mu){
  X = rnorm(n,mu,1)
  z = sqrt(n)*mean(X)
  reject = (abs(z) > qnorm(1 - alpha/2))
  return(reject)
}
power_sims <- pows_df %>%
  mutate(reject = map2_dbl(n, mu, ~rerun(100, sim_fun(n =.x, mu = .y))
                           %>% unlist %>% mean))

temp_file <- tempfile() 
download.file("http://moziru.com/images/bunny-clipart-vector-8.png", destfile = temp_file) 
rabbit <- readPNG(temp_file)

temp_file <- tempfile() 
download.file("http://pngimg.com/uploads/turtle/turtle_PNG61.png", destfile = temp_file) 
turtle <- readPNG(temp_file)

temp_file <- tempfile() 
download.file("http://pngimg.com/uploads/student/student_PNG173.png", destfile = temp_file) 
goodstudent <- readPNG(temp_file)
```

## Local Power

```{r turtle}
ggplot(power_sims, aes(x = n, y = reject, colour = Type)) +
  geom_line() + 
  geom_blank()+
  annotation_custom(rasterGrob(rabbit), xmin=750, xmax= 900, ymin=0.2, ymax=0.3) +
  annotation_custom(rasterGrob(goodstudent), xmin=750, xmax= 900, ymin=0.6, ymax=0.7) +
  annotation_custom(rasterGrob(turtle), xmin=750, xmax= 900, ymin=0.85, ymax=0.95) 
```

## Comments
For rabbit, $\mu = 2/n^{0.6}$
$$\begin{align}
power &= 1-\Phi(1.96 - 2n^{-0.1}) + \Phi(-1.96 - 2n^{-0.1})\\
& \rightarrow 1- \Phi(1.96) + \Phi(-1.96) = 0.05
\end{align} 
$$

For good students, $\mu = 2/n^{0.5}$,

$$
power = 1 - \Phi(1.96 - 2) + \Phi(-1.96 - 2) \approx 0.5
$$

For turtle, $\mu = 2/n^{0.4}$,
$$
power = 1 - \Phi(1.96 -2n^{0.1} ) + \Phi(-1.96 - 2n^{0.1})  \rightarrow 1
$$

## Final Comments
- When $\beta$ follows a rabbit sequence, the distribution of post-test estimator converges slowly to restricted one (which is incorrect), power converges to 0.05 at a very slow rate. 

\[ t = \frac{\hat{\beta}}{\sqrt{\frac{1}{n}\hat{V}_{\beta}}} 
= \frac{\sqrt{n}(\hat{\beta} - \beta)}{\sqrt{\hat{V}_{\beta}}} + \frac{\sqrt{n}\beta}{\sqrt{\hat{V}_{\beta}}} = Z + \sqrt{n}\frac{\beta}{\sqrt{\hat{V}_{\beta}}}\]
With $\beta = 2/n^{0.6}$,
$$t = Z + \frac{2n^{-0.1}}{\sqrt{\hat{V}_{\beta}}} \rightarrow Z$$

$$Pr( |t| >c | H_a \ is \ true ) \rightarrow Pr( |Z|>c ) = \alpha = 0.05$$
