---
title: "Homework6"
author: "Fledging RUCers"
date: "2018/3/1"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(broom)
library(tidyverse)
```

## Analytical Results

\[\begin{aligned}
power&=Pr(Reject H_0|H_1)\\
&=Pr(|t|>c|\mu=\mu_0)\\
&=Pr(\frac{\overline{x}-0}{\sqrt{\frac{1}{n}}}>c|\mu=\mu_0)+Pr(\frac{\overline{x}-0}{\sqrt{\frac{1}{n}}}<-c|\mu=\mu_0)\\
&=Pr(\frac{\overline{x}-\mu_0}{\sqrt{\frac{1}{n}}}>\frac{c\sqrt{\frac{1}{n}}-\mu_0}{\sqrt{\frac{1}{n}}}|\mu=\mu_0)+Pr(\frac{\overline{x}-\mu_0}{\sqrt{\frac{1}{n}}}<\frac{-c\sqrt{\frac{1}{n}}-\mu_0}{\sqrt{\frac{1}{n}}}|\mu=\mu_0)\\
&=1-\Phi(c-\frac{\mu_0}{\sqrt{\frac{1}{n}}})+\Phi(-c-\frac{\mu_0}{\sqrt{\frac{1}{n}}})
\end{aligned}\]

```{r simulation results}
library(tidyverse)
library(broom)
set.seed(2018)
test=function(x,mu,df){
  t=(mean(x)-0)/(sqrt(1/df))
  gauge=qnorm(0.025,lower.tail=FALSE)
  res=ifelse(abs(t)>gauge,yes=1,no=0)
  return(res)
}
simu=function(mu,n){
  n=rep(n,100)
  res=map2(n,mu,~rnorm(.x,mean=.y)%>%test(mu=.y,df=.x))%>%unlist()
  return(sum(res)/100)  
}
mu=seq(-3,3,by=1/5)
n=(2:30)^2
input=expand.grid(n,mu)
colnames(input)=c("n","mu")
res=mutate(input,prob=map2(input$mu,input$n,~simu(mu=.x,n=.y)))%>%
  unnest(prob)
```

## Question 1

```{r}
ggplot(data=res)+
  geom_tile(aes(x=factor(mu),y=factor(n),fill=prob))
```

##Question 2

```{r}
res1=res%>%
  mutate(true=1-pnorm(qnorm(0.975)-sqrt(n)*mu)+pnorm(qnorm(0.025)-sqrt(n)*mu))
res1%>%
  filter(n %in% c(4,25,100,361)) %>%
  ggplot(aes(color =factor(n))) +
  geom_point(aes(x=mu,y=prob)) +
  geom_line(aes(x=mu,y=true)) 
```

##Question3

```{r}
res1%>% 
  filter(mu %in% seq(-3,3,by = 1/5)[c(16:20)]) %>%
  ggplot(aes(colour = factor(mu))) +
  geom_point(aes(x=n,y=prob)) +
  geom_line(aes(x=n,y=true)) 
```

##Question5

```{r}
n=500
myfunction = function(n,u=1){
  return(1- pnorm(qnorm(0.975)-sqrt(n)*u) + pnorm(qnorm(0.025) - sqrt(n)*u))
}
myfunction(n=n)
df = data.frame(n = 1:500)
df$power = myfunction(n = df$n)
df$nstar = abs(df$power - 0.9)
which(df$nstar == min(df$nstar))
```
