---
title: "Computing Assignment VI"
author: "OMG"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Analytical results: Power

The power is one minus the probability of a type II error.

$\pi(\beta) = Pr(Reject H_0|H_1 = true) = Pr(|t|>c|\mu)$  

$=Pr(t>c|\mu)+Pr(t<-c|\mu)$

$=1-Pr(\frac{\bar x}{\frac{1}{\sqrt n}}<c|\mu) + Pr(\frac{\bar x}{\frac{1}{\sqrt n}}<-c|\mu)$

$=1-\phi(\frac{c*\frac{1}{\sqrt n}-\mu}{\sqrt\frac{1}{n}}) + \phi(\frac{-c*\frac{1}{\sqrt n}-\mu}{\sqrt\frac{1}{n}})$

$=1-\phi({c-\mu*\sqrt n}) + \phi({-c-\mu*\sqrt n})$

##Heatmap of Rejection Rates

```{r, include=FALSE}
library(knitr)
library(tidyverse)
library(dplyr)
library(gridExtra)
library(kableExtra)
#estimate beta of the linear projection model

simulate<-function (simul,n,mu){
  X <- rnorm(n=n)+mu
  mu_o=0
  crit_x_left=qnorm(0.025, mean = 0, sd = 1, lower.tail = TRUE)*sqrt(1/n)+ mu_o
  crit_x_right=-qnorm(0.025, mean = 0, sd = 1, lower.tail = TRUE)*sqrt(1/n)+ mu_o
  x_hat=mean(X)
  reject=0
  if ( (x_hat>crit_x_right) | (x_hat<crit_x_left)){
  reject=1
  }
  
  dfout=data.frame(simul=simul, n=n, mu=mu, xhat=x_hat,left=crit_x_left, right=crit_x_right, reject=reject )
  
  return(dfout)
  
}

#generate grid
S <- 100
n_pow <- (6:35)^2
mus=c(seq(from = -0.47, to = 0.47, by = 0.94/30))
mus_select=sample(mus,4,replace=F)


params_gr <- expand.grid(1:S, size=n_pow, betas=mus)
res_gr <- mapply(simulate, simul=1:S ,n=params_gr$size, mu=params_gr$betas, SIMPLIFY = FALSE) %>%
  bind_rows() %>%
  as_data_frame()

stats_coef <- res_gr %>%
  group_by(mu,n) %>%
  mutate(power=pnorm((right-mu)*sqrt(n),lower.tail = FALSE)+pnorm((left-mu)*sqrt(n))) %>%
  #mutate(power=1-pnorm((-qnorm(0.025)-mu)*sqrt(n))+pnorm((qnorm(0.025)-mu)*sqrt(n))) %>%
  summarise(rrate=mean(reject), power_theo=mean(power)) 

kable(stats_coef, digits = 3, caption = "Table 1: OLS Estimates")

# heatmap
PlotA <- ggplot(stats_coef,aes(x=factor(mu), y=factor(n), fill=rrate))+
  geom_tile()+xlab("mu")+ylab("n")+labs(fill="Rejection rate")+
  scale_x_discrete(labels = function(x) round(as.numeric(x), digits=2))+
  scale_fill_continuous(guide = guide_legend(title="Average_reject",title.postion = "bottom"))+
  theme(axis.text.x=element_text(color=c("black","transparent","transparent","transparent","transparent")))

PlotB <- ggplot(stats_coef,aes(x=factor(mu), y=factor(n), fill=power_theo))+
  geom_tile()+xlab("mu")+ylab("n")+labs(fill="Power")+
  scale_x_discrete(labels = function(x) round(as.numeric(x), digits=2))+
  scale_fill_continuous(guide = guide_legend(title="Analytical Power",title.postion = "bottom"))+
  theme(axis.text.x=element_text(color=c("black","transparent","transparent","transparent","transparent")))
```

```{r}
grid.arrange(PlotA, PlotB, nrow = 1)
```

The rejection rates closely mirror the power, as expected.

##Power Curves by n

```{r}
#2. Pick a few n values
stats_coef <- res_gr %>%
  group_by(mu,n) %>%
  mutate(power=pnorm((right-mu)*sqrt(n),lower.tail = FALSE)+pnorm((left-mu)*sqrt(n))) %>%
  #mutate(power=2-pnorm((-qnorm(0.025)-mu)*sqrt(n))-pnorm((-qnorm(0.025)+mu)*sqrt(n))) %>%
  filter(n==c(49,196,256,729,1225)) %>%
  summarise(rrate=mean(reject),power_theo=mean(power)) 
  
PlotC <- ggplot(stats_coef)+
  geom_point(aes(x=mu,y=rrate,colour=factor(n)))+
  geom_line(aes(x=mu,y=power_theo,colour=factor(n)))+
  xlab("mu")+ylab("rejection rate")+labs(colour="Analytical power by n")
plot(PlotC)
```

As n increases, we see the power curve improving so that for the same $\mu$ we now have a higher power.

##Power curves by mu

```{r}
#3. Pick a few mu values

stats_coef <- res_gr %>%
  filter(mu==mus_select) %>%
  mutate(mu=round(mu, digits = 2))%>%
  group_by(mu,n) %>%
  mutate(power=pnorm((right-mu)*sqrt(n),lower.tail = FALSE)+pnorm((left-mu)*sqrt(n))) %>%
  #mutate(power=2-pnorm((-qnorm(0.025)-mu)*sqrt(n))-pnorm((-qnorm(0.025)+mu)*sqrt(n))) %>%
  summarise(rrate=mean(reject),power_theo=mean(power)) 

PlotD <- ggplot(stats_coef)+
  geom_point(aes(x=n,y=rrate,colour=factor(mu)))+
  geom_line(aes(x=n,y=power_theo,colour=factor(mu)))+
  xlab("n")+ylab("rejection rate")+labs(colour="Analytial power by mu")
plot(PlotD)
```

The rejection rate, and the power, increases as n increases.  This happens faster for $\mu$ further from 0.

## Student questions
When $H_0:\mu=0$, $\mu=1$, and $n=500$, the power is 1.  
For $\mu=1$, we would need $n=11$ for a power of 0.90.  

```{r, include=FALSE}

#quest 5 [part a]
mu_o=0
x_left=qnorm(0.025, mean = 0, sd = 1, lower.tail = TRUE)*sqrt(1/500)+ mu_o
x_right=-qnorm(0.025, mean = 0, sd = 1, lower.tail = TRUE)*sqrt(1/500)+ mu_o
power=pnorm((x_right-1)*sqrt(500),lower.tail = FALSE)+pnorm((x_left-1)*sqrt(500))

#quest 5 [part b]
simulate2<-function (simul,n,mu){
  X <- rnorm(n=n)+mu
  mu_o=0
  crit_x_left=qnorm(0.025, mean = 0, sd = 1, lower.tail = TRUE)*sqrt(1/n)+ mu_o
  crit_x_right=-qnorm(0.025, mean = 0, sd = 1, lower.tail = TRUE)*sqrt(1/n)+ mu_o
  x_hat=mean(X)
  reject=0
  if ( (x_hat>crit_x_right) | (x_hat<crit_x_left)){
    reject=1
  }
  
  dfout=data.frame(simul=simul, n=n, mu=mu, xhat=x_hat,left=crit_x_left, right=crit_x_right, reject=reject )
  
  return(dfout)
  
}


S <- 100
res_gr2 <- mapply(simulate2, simul=1:S ,n=c(5:15), mu=1, SIMPLIFY = FALSE) %>%
  bind_rows() %>%
  as_data_frame()
```

```{r, message=FALSE, warning=FALSE}
stats_coef2 <- res_gr2 %>%
  group_by(mu,n) %>%
  mutate(power=pnorm((right-mu)*sqrt(n),lower.tail = FALSE)+pnorm((left-mu)*sqrt(n))) %>%
  #mutate(power=2-pnorm((-qnorm(0.025)-mu)*sqrt(n))-pnorm((-qnorm(0.025)+mu)*sqrt(n))) %>%
  summarise(rrate=mean(reject), power_theo=mean(power)) 

kable(stats_coef2, digits = 3, caption = "Table 1: power table", format = "html")%>%
  kable_styling(font_size = 22)%>%
  column_spec(2, width = "2em")%>%
  column_spec(3, width = "4em")

```

## The Turtle and the Hare: Logic

$$t = \frac{\bar x-\mu_o}{\frac{1}{\sqrt n}} = \frac{\bar x-\mu_a}{\frac{1}{\sqrt n}} + \frac{\mu_a -\mu_o}{\frac{1}{\sqrt n}} \xrightarrow[]{d} ~N(0,1)+\frac{\mu_a}{\frac{1}{\sqrt n}}$$ 

For the good student, $\frac{\mu_a}{\frac{1}{\sqrt n}} = \frac{\frac{2}{\sqrt n}}{\frac{1}{\sqrt n}} = 2$

For the hare, $\frac{\mu_a}{\frac{1}{\sqrt n}} = \frac{\frac{2}{n^{0.6}}}{\frac{1}{\sqrt n}} = \frac{2}{n^{0.1}} \xrightarrow[]{p} 0$

For the turtle, $\frac{\mu_a}{\frac{1}{\sqrt n}} = \frac{\frac{2}{n^{0.4}}}{\frac{1}{\sqrt n}} = 2*n^{0.1} \xrightarrow[]{p} \infty$

## The Turtle and the Hare

```{r}
#OPTIONAL
###good student
#mu=0+2/(n^0.5)
###turtle
#mu=0+2/(n^0.4)
###rabbit
#mu=0+2/(n^0.6)

sim_fo<-function (n,mu){
  X <- rnorm(n=n)+mu
  mu_o=0
  crit_x_left=qnorm(0.025, mean = 0, sd = 1, lower.tail = TRUE)*sqrt(1/n)+ mu_o
  crit_x_right=-qnorm(0.025, mean = 0, sd = 1, lower.tail = TRUE)*sqrt(1/n)+ mu_o
  x_hat=mean(X)
  reject=0
  if ( (x_hat>crit_x_right) | (x_hat<crit_x_left)){
    reject=1
  }
  
  dfout=reject
  
  return(dfout)
  
}
n_pow <- (2:30)^2
pows_df <- data_frame(Type= c("Turtle", "Good student", "Rabbit"),
                      power=c(0.4, 0.5, 0.6)) %>%
  mutate(mu = map(power, ~data_frame(n=n_pow, mu=2/n_pow^.))) %>%
  unnest(mu)

pows_sims <- pows_df %>%
  mutate(reject=map2_dbl(n, mu, ~rerun(100, sim_fo(n=.x, mu=.y)) %>% unlist %>% mean ))

library(png)
library(grid)
library(gridExtra)
library(jpeg)
## function readPNG() does not accept an URL, so we need to download first:
temp_file <- tempfile()
download.file("http://moziru.com/images/bunny-clipart-vector-8.png", destfile = temp_file)
# if it doesnt work, try one of:
download.file("http://moziru.com/images/bunny-clipart-vector-8.png", destfile = temp_file, mode="wb")
#download.file("http://moziru.com/images/bunny-clipart-vector-8.png", destfile = temp_file, method="wget")
img <- readPNG(temp_file)
temp_file2<-tempfile()
download.file("https://www.usmagazine.com/wp-content/uploads/finding-dory-baby-turtle-bc12b198-fefc-44d6-8be5-07ddcef85b04.jpg", destfile = temp_file2,mode="wb")
img2 <- readJPEG(temp_file2)
## How to add a rabbit on an empty plot (geom_blank())

ggplot(pows_sims, aes(x=n,y=reject,colour=Type)) +
  geom_point()+
  annotation_custom(rasterGrob(img), xmin=750, xmax= 900, ymin=0.2, ymax=0.3)+
  annotation_custom(rasterGrob(img2), xmin=750, xmax= 900, ymin=0.85, ymax=0.95)
```

