---
title: "Assignment 6"
author: "OZONE"
date: "February 28, 2018"
output: beamer_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(broom)
library(tidyr)
library(tidyverse)
library(ggplot2)
```

## Analytical Results

The power of a test is the probability of rejecting null hypothesis when the null is false. Hence, given $\mu$ and sample size, we could calculate this probability.
$$\begin{aligned}
power =& Pr(reject\ H_0|H_0\ not\ true)\\
=& Pr(\frac{\bar{x}-0}{\sqrt{\frac{1}{n}}}>c|\mu=\mu^{*}) + Pr(\frac{\bar{x}-0}{\sqrt{\frac{1}{n}}}<-c|\mu=\mu^{*})\\
=& Pr(\frac{\bar{x}-\mu^{*}}{\sqrt{\frac{1}{n}}}>\frac{c\sqrt{\frac{1}{n}}-\mu^{*}}{\sqrt{\frac{1}{n}}}|\mu=\mu^{*}) + \\
&Pr(\frac{\bar{x}-\mu^{*}}{\sqrt{\frac{1}{n}}}<\frac{-c\sqrt{\frac{1}{n}}-\mu^{*}}{\sqrt{\frac{1}{n}}}|\mu=\mu^{*})\\
=& 1-\Phi(c-\frac{\mu^*}{\sqrt{\frac{1}{n}}}) + \Phi(-c-\frac{\mu^*}{\sqrt{\frac{1}{n}}})
\end{aligned}$$


```{r simulation results}
set.seed(123)
sim_dat=function(n=10,u=0){
 dat=data.frame(x=rnorm(n,u,1))
 result=data.frame( z  = mean(dat$x)/sqrt(1/n)) %>%
 mutate(test= ifelse (z < qnorm(.025) | z > qnorm(0.975), 1, 0)) #1=reject;0=not reject
 return(result)
}

grid=expand.grid(n=(2:31)^2, u=seq(-1,1,by=2/29))%>%
  mutate(test=map2(n,u,~rerun(100,sim_dat(n=.x,u=.y))%>% bind_rows))%>%
  unnest(test)
Rejection= grid %>% group_by(n,u)%>% summarise(ave_prob=mean(test))

```

##Question 1
```{r}
ggplot(aes(factor(n),u),data=Rejection)+
  geom_tile(aes(fill=ave_prob))
```

##Question 2
```{r}
Rejection = Rejection %>% mutate(any = 1- pnorm(qnorm(0.975)-sqrt(n)*u) + pnorm(qnorm(0.025) - sqrt(n)*u))
Rejection %>% filter(n %in% c(9,36,121,400)) %>%
  ggplot(aes(colour = factor(n))) +
  geom_point(aes(x=u,y=ave_prob)) +
  geom_line(aes(x=u,y=any)) 
```
- Note: Dots are simulated data. Lines are analytical curves.

##Question3
```{r}
Rejection %>% filter(u %in% seq(-1,1,by = 2/29)[c(16:20)]) %>%
ggplot(aes(colour = factor(u))) +
  geom_point(aes(x=n,y=ave_prob)) +
  geom_line(aes(x=n,y=any)) 
```
- Note: Dots are simulated data. Lines are analytical curves.

##Question5
- When n=500, I expect to detect a difference of 1 in my outcome variable with almost 100% confidence.
- When the sample size is 11, we could detect a difference of 1 with 90% confidence.
```{r}
n=500
myfunction = function(n,u=1){
  return(1- pnorm(qnorm(0.975)-sqrt(n)*u) + pnorm(qnorm(0.025) - sqrt(n)*u))
}
myfunction(n=n)
df = data.frame(n = 1:500)
df$power = myfunction(n = df$n)
df$nstar = abs(df$power - 0.9)
which(df$nstar == min(df$nstar))
```




