---
title: "PSX5"
author: "Surplus Value"
date: "2/22/2018"
output: 
beamer_presentation:
latex_engine: xelatex
---


##Start
```{r chunk_name, echo=FALSE}
library(sandwich)
library(lmtest)
library(broom)
library(tidyverse)
library(MASS)

mu <- c(0,0)
Sigma <- as.matrix(rbind(c(1,0.7),c(0.7,1)))
```

##Function
```{r , echo=FALSE}
simul <- function(n=n, beta=beta) {
  alpha <- rep(0.2,n)
  beta  <- rep(beta,n)
  e <- rnorm(n, mean=0, sd=1)
  X <- mvrnorm(n,mu,Sigma)
  X1<- as.vector(X[,1])
  X2<- as.vector(X[,2])
  y <- alpha*X1+beta*X2 + e
  df <- data_frame(y,X1,X2)
  reg <-lm(y~X1 + X2-1, data=df)
  alphaHat <- coeftest(reg)[1,1]
  betaHatPvalue <- coeftest(reg)[2,4]
  Conf <- 0.2 > confint(reg)["X1",1] & 0.2 < confint(reg)["X1",2]
  regres <-lm(y~X1-1, data=df)
  alphaHatRes <- coeftest(regres)[1,1]
  ConfRes <- 0.2 > confint(regres)["X1",1] & 0.2 < confint(regres)["X1",2]
  
  
  if (coef(summary(reg))[2,1] < .05) {
    reg_everyday <- reg
  }
  

  if (coef(summary(reg))[2,1] > .05) {
    reg_everyday <- regres
  }
  
  Confeveryday <- 0.2 > confint(reg_everyday)["X1",1] & 0.2 < confint(reg_everyday)["X1",2]
  
  
  rs<- c(alphaHat,betaHatPvalue,Conf,alphaHatRes,ConfRes,coeftest(reg_everyday)[1,1],Confeveryday)
  return(rs)
}
```

#1
```{r , echo=FALSE}
S <- 2000
params_gr <-expand.grid(1:S, n=c(50,100,150,200),  beta=c(0,.16,.24,.5))
res_gr <-mapply(simul, n=params_gr$n, beta=params_gr$beta, SIMPLIFY = FALSE)
frame <- data.frame(matrix(unlist(res_gr), ncol=7, byrow = S*16))
dfchunk <- split(frame, factor(sort(rank(row.names(frame))%%4)))
df0 <- as.data.frame(dfchunk["0"])
df16 <- as.data.frame(dfchunk["1"])
df24 <- as.data.frame(dfchunk["2"])
df5 <- as.data.frame(dfchunk["3"])

head(df0)
head(df16)
head(df24)
head(df5)

```

#2
```{r , echo=FALSE}
var(df0[,1]) #Variance of unrestricted
var(df0[,4]) #Variance of restricted
var(df0[,6]) #variance of every day



```

#####Since X2 does not explain anything  in the model, the variances of the three estimators are very close to each other.

#3
```{r , echo=FALSE}
Bias16 <- .16*.7
Bias24 <- .25*.7
Bias5 <-  .5*.7
#Bias of the restricted estimators
Ealpha16_r <- mean(df16[,4]-0.2)
Ealpha24_r <- mean(df24[,4]-0.2)
Ealpha5_r<- mean(df5[,4]-0.2)

#Bias of the unrestricted estimators
Ealpha16 <- mean(df16[,1]-0.2)
Ealpha24 <- mean(df24[,1]-0.2)
Ealpha5 <- mean(df5[,1]-0.2)

#Bias for the every day estimators
Ealpha16_ed <- mean(df16[,6]-0.2)
Ealpha24_ed <- mean(df24[,6]-0.2)
Ealpha5_ed<- mean(df5[,6]-0.2)





```

```{r}


#The unrestricted model:

Ealpha16
Ealpha24
Ealpha5


#The unrestricted model:

Ealpha16_r
Ealpha24_r
Ealpha5_r

#The everyday model:

Ealpha16_ed
Ealpha24_ed
Ealpha5_ed

#The OVB:
Bias16 
Bias24 
Bias5

```

#####In general the unrestricted model gives as expected almost zero bias. The restricted and everyday biases, and are very close to the value that is derived from the omited variable bias formula. 


#4
```{r , echo=FALSE}

#By summing variables =0,1 we get a number of times we reject
NumAl0 <- sum(df0[,3])
NumAl0R <- sum(df0[,5])
NumAl0ED <- sum(df0[,7])
NumAl16 <- sum(df16[,3])
NumAl16R <- sum(df16[,5])
NumAl16ED <- sum(df16[,7])
NumAl24 <- sum(df24[,3])
NumAl24R <- sum(df24[,5])
NumAl24ED <- sum(df24[,7])
NumAl5 <- sum(df5[,3])
NumAl5R <- sum(df5[,5])
NumAl5ED <- sum(df5[,7])
```

```{r}
##Unrestricted Model
NumAl0
NumAl16
NumAl24
NumAl5

##Restricted Model

NumAl0R
NumAl16R
NumAl24R
NumAl5R

##EveryDay Model

NumAl0ED
NumAl16ED
NumAl24ED
NumAl5ED


```

#####As expected the number of times that the true value within the CI are very close to each other, no matter how high the beta of X2 was. Under the restricted model and the everyday, the greater the beta, the smaller number of times that the true value was in the CI. The number drops much faster for the restricted model. 

#5
```{r , echo=FALSE}

library(ggplot2)

dfchunk0 <- split(df0, factor(sort(rank(row.names(df0))%%4)))
dfN50_B0 <- as.data.frame(dfchunk0["0"])
dfN100_B0 <- as.data.frame(dfchunk0["1"])
dfN150_B0 <- as.data.frame(dfchunk0["2"])
dfN200_B0 <- as.data.frame(dfchunk0["3"])

dfchunk16 <- split(df16, factor(sort(rank(row.names(df16))%%4)))
dfN50_B16 <- as.data.frame(dfchunk16["0"])
dfN100_B16 <- as.data.frame(dfchunk16["1"])
dfN150_B16 <- as.data.frame(dfchunk16["2"])
dfN200_B16 <- as.data.frame(dfchunk16["3"])

dfchunk24 <- split(df24, factor(sort(rank(row.names(df24))%%4)))
dfN50_B24 <- as.data.frame(dfchunk24["0"])
dfN100_B24 <- as.data.frame(dfchunk24["1"])
dfN150_B24 <- as.data.frame(dfchunk24["2"])
dfN200_B24 <- as.data.frame(dfchunk24["3"])

dfchunk5 <- split(df5, factor(sort(rank(row.names(df5))%%4)))
dfN50_B5 <- as.data.frame(dfchunk5["0"])
dfN100_B5 <- as.data.frame(dfchunk5["1"])
dfN150_B5 <- as.data.frame(dfchunk5["2"])
dfN200_B5 <- as.data.frame(dfchunk5["3"])

#Unrestricted Data by sample size and by beta
stdUn_N50_B0 <- data.frame(x1=50^.5 * (dfN50_B0[,1]-0.2), b=c(0), n=c(50), line=c("Unrestricted"))
stdUn_N100_B0 <- data.frame(x1=100^.5 * (dfN100_B0[,1]-0.2), b=c(0), n=c(100), line=c("Unrestricted"))
stdUn_N150_B0 <- data.frame(x1=150^.5 * (dfN150_B0[,1]-0.2),b=c(0), n=c(150), line=c("Unrestricted"))
stdUn_B200_B0 <- data.frame(x1=200^.5 * (dfN200_B0[,1]-0.2),b=c(0), n=c(200), line=c("Unrestricted"))

stdUn_N50_B16 <- data.frame(x1=50^.5 * (dfN50_B16[,1]-0.2), b=c(0.16), n=c(50), line=c("Unrestricted"))
stdUn_N100_B16 <-data.frame(x1= 100^.5 * (dfN100_B16[,1]-0.2), b=c(0.16), n=c(100), line=c("Unrestricted"))
stdUn_N150_B16 <- data.frame(x1=150^.5 * (dfN150_B16[,1]-0.2), b=c(0.16), n=c(150), line=c("Unrestricted"))
stdUn_B200_B16 <- data.frame(x1=200^.5 * (dfN200_B16[,1]-0.2), b=c(0.16), n=c(200), line=c("Unrestricted"))

stdUn_N50_B24 <- data.frame(x1=50^.5 * (dfN50_B24[,1]-0.2), b=c(0.24), n=c(50), line=c("Unrestricted"))
stdUn_N100_B24 <-data.frame(x1= 100^.5 * (dfN100_B24[,1]-0.2),b=c(0.24), n=c(100), line=c("Unrestricted"))
stdUn_N150_B24 <- data.frame(x1=150^.5 * (dfN150_B24[,1]-0.2),b=c(0.24), n=c(150), line=c("Unrestricted"))
stdUn_B200_B24 <- data.frame(x1=200^.5 * (dfN200_B24[,1]-0.2),b=c(0.24), n=c(200), line=c("Unrestricted"))

stdUn_N50_B5 <- data.frame(x1=50^.5 * (dfN50_B5[,1]-0.2), b=c(0.5), n=c(50), line=c("Unrestricted"))
stdUn_N100_B5 <- data.frame(x1=100^.5 * (dfN100_B5[,1]-0.2), b=c(0.5), n=c(100), line=c("Unrestricted"))
stdUn_N150_B5 <- data.frame(x1=150^.5 * (dfN150_B5[,1]-0.2), b=c(0.5), n=c(150), line=c("Unrestricted"))
stdUn_B200_B5 <- data.frame(x1=200^.5 * (dfN200_B5[,1]-0.2), b=c(0.5), n=c(200), line=c("Unrestricted"))


#All unrestricted data combined
hyperdata_un= rbind(stdUn_N50_B0,stdUn_N100_B0,stdUn_N150_B0,stdUn_B200_B0,
                    stdUn_N50_B16,stdUn_N100_B16,stdUn_N150_B16,stdUn_B200_B16,
                    stdUn_N50_B24,stdUn_N100_B24,stdUn_N150_B24,stdUn_B200_B24,
                    stdUn_N50_B5,stdUn_N100_B5,stdUn_N150_B5,stdUn_B200_B5)




#Restricted Data by sample size and by beta
stdres_N50_B0 <- data.frame(x1=50^.5 * (dfN50_B0[,4]-0.2), b=c(0), n=c(50), line=c("Restricted"))
stdres_N100_B0 <- data.frame(x1=100^.5 * (dfN100_B0[,4]-0.2), b=c(0), n=c(100), line=c("Restricted"))
stdres_N150_B0 <- data.frame(x1=150^.5 * (dfN150_B0[,4]-0.2),b=c(0), n=c(150), line=c("Restricted"))
stdres_B200_B0 <- data.frame(x1=200^.5 * (dfN200_B0[,4]-0.2),b=c(0), n=c(200), line=c("Restricted"))

stdres_N50_B16 <- data.frame(x1=50^.5 * (dfN50_B16[,4]-0.2), b=c(0.16), n=c(50), line=c("Restricted"))
stdres_N100_B16 <-data.frame(x1= 100^.5 * (dfN100_B16[,4]-0.2), b=c(0.16), n=c(100), line=c("Restricted"))
stdres_N150_B16 <- data.frame(x1=150^.5 * (dfN150_B16[,4]-0.2), b=c(0.16), n=c(150), line=c("Restricted"))
stdres_B200_B16 <- data.frame(x1=200^.5 * (dfN200_B16[,4]-0.2), b=c(0.16), n=c(200), line=c("Restricted"))

stdres_N50_B24 <- data.frame(x1=50^.5 * (dfN50_B24[,4]-0.2), b=c(0.24), n=c(50), line=c("Restricted"))
stdres_N100_B24 <-data.frame(x1= 100^.5 * (dfN100_B24[,4]-0.2),b=c(0.24), n=c(100), line=c("Restricted"))
stdres_N150_B24 <- data.frame(x1=150^.5 * (dfN150_B24[,4]-0.2),b=c(0.24), n=c(150), line=c("Restricted"))
stdres_B200_B24 <- data.frame(x1=200^.5 * (dfN200_B24[,4]-0.2),b=c(0.24), n=c(200), line=c("Restricted"))

stdres_N50_B5 <- data.frame(x1=50^.5 * (dfN50_B5[,4]-0.2), b=c(0.5), n=c(50), line=c("Restricted"))
stdres_N100_B5 <- data.frame(x1=100^.5 * (dfN100_B5[,4]-0.2), b=c(0.5), n=c(100), line=c("Restricted"))
stdres_N150_B5 <- data.frame(x1=150^.5 * (dfN150_B5[,4]-0.2), b=c(0.5), n=c(150), line=c("Restricted"))
stdres_B200_B5 <- data.frame(x1=200^.5 * (dfN200_B5[,4]-0.2), b=c(0.5), n=c(200), line=c("Restricted"))


#All Restricted data combined
hyperdata_res= rbind(stdres_N50_B0,stdres_N100_B0,stdres_N150_B0,stdres_B200_B0,
                     stdres_N50_B16,stdres_N100_B16,stdres_N150_B16,stdres_B200_B16,
                     stdres_N50_B24,stdres_N100_B24,stdres_N150_B24,stdres_B200_B24,
                     stdres_N50_B5,stdres_N100_B5,stdres_N150_B5,stdres_B200_B5)


#Every Day
stded_N50_B0 <- data.frame(x1=50^.5 * (dfN50_B0[,6]-0.2), b=c(0), n=c(50), line=c("Everyday"))
stded_N100_B0 <- data.frame(x1=100^.5 * (dfN100_B0[,6]-0.2), b=c(0), n=c(100), line=c("Everyday"))
stded_N150_B0 <- data.frame(x1=150^.5 * (dfN150_B0[,6]-0.2),b=c(0), n=c(150), line=c("Everyday"))
stded_B200_B0 <- data.frame(x1=200^.5 * (dfN200_B0[,6]-0.2),b=c(0), n=c(200), line=c("Everyday"))

stded_N50_B16 <- data.frame(x1=50^.5 * (dfN50_B16[,6]-0.2), b=c(0.16), n=c(50), line=c("Everyday"))
stded_N100_B16 <-data.frame(x1= 100^.5 * (dfN100_B16[,6]-0.2), b=c(0.16), n=c(100), line=c("Everyday"))
stded_N150_B16 <- data.frame(x1=150^.5 * (dfN150_B16[,6]-0.2), b=c(0.16), n=c(150), line=c("Everyday"))
stded_B200_B16 <- data.frame(x1=200^.5 * (dfN200_B16[,6]-0.2), b=c(0.16), n=c(200), line=c("Everyday"))

stded_N50_B24 <- data.frame(x1=50^.5 * (dfN50_B24[,6]-0.2), b=c(0.24), n=c(50), line=c("Everyday"))
stded_N100_B24 <-data.frame(x1= 100^.5 * (dfN100_B24[,6]-0.2),b=c(0.24), n=c(100), line=c("Everyday"))
stded_N150_B24 <- data.frame(x1=150^.5 * (dfN150_B24[,6]-0.2),b=c(0.24), n=c(150), line=c("Everyday"))
stded_B200_B24 <- data.frame(x1=200^.5 * (dfN200_B24[,6]-0.2),b=c(0.24), n=c(200), line=c("Everyday"))

stded_N50_B5 <- data.frame(x1=50^.5 * (dfN50_B5[,6]-0.2), b=c(0.5), n=c(50), line=c("Everyday"))
stded_N100_B5 <- data.frame(x1=100^.5 * (dfN100_B5[,6]-0.2), b=c(0.5), n=c(100), line=c("Everyday"))
stded_N150_B5 <- data.frame(x1=150^.5 * (dfN150_B5[,6]-0.2), b=c(0.5), n=c(150), line=c("Everyday"))
stded_B200_B5 <- data.frame(x1=200^.5 * (dfN200_B5[,6]-0.2), b=c(0.5), n=c(200), line=c("Everyday"))

```


```{r}
#All Every Day data combined
hyperdata_ed= rbind(stded_N50_B0,stded_N100_B0,stded_N150_B0,stded_B200_B0,
                    stded_N50_B16,stded_N100_B16,stded_N150_B16,stded_B200_B16,
                    stded_N50_B24,stded_N100_B24,stded_N150_B24,stded_B200_B24,
                    stded_N50_B5,stded_N100_B5,stded_N150_B5,stded_B200_B5)

ggplot(aes(x=x1, fill="Restricted"), data=hyperdata_res)+ 
  geom_density(alpha=0.5)+
  facet_grid(n~b)

ggplot(aes(x=x1, fill="Unrestricted"), data=hyperdata_un)+ 
  geom_density(alpha=0.5)+
  facet_grid(n~b)

dat2 <- data.frame(rbind(hyperdata_un, hyperdata_res ))

ggplot(aes(x=x1, fill=line), data=dat2)+ 
  geom_density(alpha=0.5)+
  facet_grid(n~b)
```

#####The more X2 explains the variation of Y, the further the distribution of the estimator is from the true value.

#6
```{r}

#all three estimators

dat1 <- data.frame(rbind(hyperdata_un, hyperdata_res,hyperdata_ed ))

ggplot(aes(x=x1, fill=line), data=dat1)+ 
  geom_density(alpha=0.5)+
  facet_grid(n~b)

```



