---
title: "Computing Assignment VII"
author: "Ozone"
date: "March 9, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
knitr::opts_knit$set(root.dir = "~/Downloads/240A_Assign_7_data")
library(stargazer)
library(AER)
```

##Q1
```{r, results='asis'}
mydata = read.csv("RSdata.csv")
OLS1   = lm(ln_qs ~ ln_fp + ln_w + trendsp1 + trendsp2 + trendsp3, data = mydata)
OLS2   = lm(ln_qs ~ ln_fp + trendsp1 + trendsp2 + trendsp3, data = mydata)
OLS3   = lm(ln_qd ~ ln_sp + trendsp1 + trendsp2 + trendsp3, data = mydata)

stargazer(OLS1,OLS2,OLS3,header = F, column.labels = c("Supply","Supply","Demand"),dep.var.labels = c("",""),title = "Q1")
```

Authors did not include log yield shock when estimate supply elasticity. 

##Q2
```{r,results='asis'}
IV1   = ivreg(ln_qs ~ ln_fp + ln_w + trendsp1 + trendsp2 + trendsp3|.-ln_fp + lag_ln_w, data = mydata)
IV2   = ivreg(ln_qd ~ ln_sp + trendsp1 + trendsp2 + trendsp3|.-ln_sp + ln_w, data = mydata)
stargazer(IV1,IV2,header = F, column.labels = c("Supply","Demand"),dep.var.labels = c("",""),title = "Q2")
```

##Q3

We reject the first test and do not reject the second test.
```{r,results='asis'}
H1 = (IV1$coefficients[2] - OLS2$coefficients[2])^2*(summary(IV1)[["coefficients"]][2,2]^2 - summary(OLS2)[["coefficients"]][2,2]^2)^(-1)
pchisq(H1,1)

H2 = (IV1$coefficients[2] - OLS1$coefficients[2])^2*(summary(IV1)[["coefficients"]][2,2]^2 - summary(OLS1)[["coefficients"]][2,2]^2)^(-1)
pchisq(H2,1)
```

##Q4
```{r, results='asis'}
mydata$ln_area = log(mydata$area)
OLS4  = lm(ln_area ~ ln_fp + ln_w + trendsp1 + trendsp2 + trendsp3, data = mydata)
OLS5  = lm(ln_area ~ ln_fp + trendsp1 + trendsp2 + trendsp3, data = mydata)
IV3   = ivreg(ln_area ~ ln_fp + ln_w + trendsp1 + trendsp2 + trendsp3|.-ln_fp + lag_ln_w, data = mydata)
stargazer(OLS4,OLS5,IV3,header = F,dep.var.labels = c("","",""),title = "Q4")
```

We observe a difference around 20% between the total quantity model and area model, which implies part of the change in total supply is from yield adjustment.

##Q5

From the estimations in previous questions, we notice discrepancies after adding 7 years data.

##Q6

The data used by author is in long form and not easy to understand. Also it is not easy to manupalate. 
