---
title: "Computing Assignment VII"
author: "OMG"
date: "March 9, 2018"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
library(tidyverse)
library(AER)
library(broom)
library(MASS)
library(tidyverse)
library(knitr)
setwd("..")
dta2 <- read_csv("data/RSdata.csv")
#dta <- read_csv("https://github.com/ucdavis/ARE_ECON_240A_2018/tree/master/Computing_Assignment_VII/data/RSdata.csv")
dta<-dta2%>%filter(year<=2007)
```

## Simple OLS and IV

```{r include=FALSE}

#lm(ln_qs~ln_fp+ln_w+trendsp1+trendsp2,dta)
#lm(ln_qs~ln_fp+ln_w+trendsp1+trendsp3,dta)
##supply
OLSS=lm(ln_qs~ln_fp+trendsp1+trendsp2+trendsp3,dta)
OLSSw=lm(ln_qs~ln_fp+ln_w+trendsp1+trendsp2+trendsp3,dta)
#a=lm(ln_qs~ln_fp+ln_w+trendsp1+trendsp2+trendsp3,dta)
##demand
OLSD=lm(ln_qd~ln_sp+trendsp1+trendsp2+trendsp3,dta)
## dropping the ln_w term gives us somewhat closer value

##2. IV
## supply model lag of ln_w as IV for ln_fp
IVS <- ivreg(ln_qs~ln_fp+ln_w+trendsp1+trendsp2+trendsp3|ln_w+trendsp1+trendsp2+trendsp3+lag_ln_w,data=dta)
IVD <- ivreg(ln_qd~ln_sp+trendsp1+trendsp2+trendsp3|trendsp1+trendsp2+trendsp3+ln_w,data=dta)

```

Estimating the supply funcion using OLS, we found that excluding ln(w) led to a result more similar to the paper.  

|OLS |demand | supply |supply, with ln(w) |
|--:|----:|---------:|----:|
| | -0.021| 0.023|0.097|
| | (0.011)| (0.027)|(0.013)|

|IV |demand | supply, with ln(w) |
|--:|----:|---------:|
| | -0.059| 0.010|
| | (0.022)|  (0.024)|


##The Hausman Test

$H_0:$ Price is exogenous  
$H_1:$ Price is endogenous

$\hat\beta^{IV} \xrightarrow{p} \beta$ under both $H_0$ and $H_1$  
$\hat\beta^{OLS} \xrightarrow{p} \beta$ under $H_0:$, is more efficient.

$H=(\hat\beta^{IV}-\hat\beta^{OLS})'(\hat {var}(\hat\beta^{IV}-\hat\beta^{OLS}))^{-1}(\hat\beta^{IV}-\hat\beta^{OLS})$
$= \frac{(\hat\beta^{IV}-\hat\beta^{OLS})^2}{\hat {var}(\hat\beta^{IV}-\hat\beta^{OLS})}$
if there is only one parameter.  

Assuming homoskedasticity:$\hat {var}(\hat\beta^{IV}-\hat\beta^{OLS})= var(\hat\beta^{IV})-var(\hat\beta^{OLS})$


##Hausman results

```{r}

# excluding lnw from OLS
#homoscedastic case
b_iv=summary(IVS)$coef[2,1]
var_b_iv=(summary(IVS)$coef[2,2])^2

b_ols=summary(OLSS)$coef[2,1]
var_b_ols=(summary(OLSS)$coef[2,2])^2

var_delta_b=var_b_iv-var_b_ols
H=((b_iv-b_ols)^2)/var_delta_b

critical=pchisq(0.95,1)

#h2 <- hausman.systemfit( fit2sls, fit3sls ) 

# including lnw in OLS
OLSS2=lm(ln_qs~ln_fp+ln_w+trendsp1+trendsp2+trendsp3,dta)
b_ols2=summary(OLSS2)$coef[2,1]
var_b_ols2=(summary(OLSS2)$coef[2,2])^2
var_delta_b2=var_b_iv-var_b_ols2
H2=((b_iv-b_ols2)^2)/var_delta_b2
```

Our Hausman results testing the OLS without ln(w) against the IV:

$$H= -44.393$$

And the results testing the OLS with ln(w) against the IV:

$$H= 0.076$$
$$c = 0.670$$  
We reject the null hypothesis that price is exogenous in the first case and fail to reject in the second.  


```{r}
###4 Repeat supply model with ln(area) as dependent variable
###ln_w is not included 
dta <- dta %>% mutate(ln_area=log(area)) 
olsarea=lm(ln_area~ln_fp+ln_w+trendsp1+trendsp2+trendsp3,dta) %>% tidy
IVarea <- ivreg(ln_area~ln_fp+ln_w+trendsp1+trendsp2+trendsp3|ln_w+trendsp1+trendsp2+trendsp3+lag_ln_w,data=dta) %>% tidy

###5 adding the 7 extra years doesn't seem to change the IV estimates for the supply and demand model. 
###However, it did seem to affect our Hausmen test results. Whether Ho: price is exogenous or not.
```

## your document
Geyi_Hi_I hope that you could see this lol...