---
title: "Computing Assignment VII"
author: "OMG"
date: "March 9, 2018"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
library(tidyverse)
library(AER)
library(broom)
library(MASS)
library(tidyverse)
library(knitr)
setwd("..")
dta2 <- read_csv("data/RSdata.csv")
#dta <- read_csv("https://github.com/ucdavis/ARE_ECON_240A_2018/tree/master/Computing_Assignment_VII/data/RSdata.csv")
dta<-dta2%>%filter(year<=2007)
```

## Simple OLS and IV

```{r echo=FALSE}

#lm(ln_qs~ln_fp+ln_w+trendsp1+trendsp2,dta)
#lm(ln_qs~ln_fp+ln_w+trendsp1+trendsp3,dta)
##supply
OLSSt=lm(ln_qs~ln_fp+trendsp1+trendsp2+trendsp3,dta)%>% tidy
tOLSS <- OLSSt[-c(1, 3, 4, 5), ]
OLSS2t=lm(ln_qs~ln_fp+ln_w+trendsp1+trendsp2+trendsp3,dta)%>% tidy
tOLSS2 <- OLSS2t[-c(1, 3, 4, 5, 6), ]
#a=lm(ln_qs~ln_fp+ln_w+trendsp1+trendsp2+trendsp3,dta)
##demand
OLSDt=lm(ln_qd~ln_sp+trendsp1+trendsp2+trendsp3,dta)%>% tidy
tOLSD <- OLSDt[-c(1, 3, 4, 5), ]
## dropping the ln_w term gives us somewhat closer value

IVSt <- ivreg(ln_qs~ln_fp+ln_w+trendsp1+trendsp2+trendsp3|ln_w+trendsp1+trendsp2+trendsp3+lag_ln_w,data=dta)%>% tidy
tIVS <- IVSt[-c(1, 3, 4, 5, 6), ]

IVDt <- ivreg(ln_qd~ln_sp+trendsp1+trendsp2+trendsp3|trendsp1+trendsp2+trendsp3+ln_w,data=dta)%>% tidy
tIVD <- IVDt[-c(1, 3, 4, 5, 6), ]


OLSIVtable <- rbind(tOLSS, tOLSS2, tOLSD, tIVS, tIVD)
rownames(OLSIVtable)= c("OLS S", "OLS S with ln(w)", "OLS D", "IV S", "IV D")

kable(OLSIVtable)


##2. IV
## supply model lag of ln_w as IV for ln_fp
#IVS <- ivreg(ln_qs~ln_fp+ln_w+trendsp1+trendsp2+trendsp3|ln_w+trendsp1+trendsp2+trendsp3+lag_ln_w,data=dta)
#IVD <- ivreg(ln_qd~ln_sp+trendsp1+trendsp2+trendsp3|trendsp1+trendsp2+trendsp3+ln_w,data=dta)

```

Estimating the supply funcion using OLS, we found that excluding ln(w) led to a result more similar to the paper.  



##The Hausman Test

$H_0:$ Price(ln_fp) is exogenous  
$H_1:$ Price is endogenous

$\hat\beta^{IV} \xrightarrow{p} \beta$ under both $H_0$ and $H_1$  
$\hat\beta^{OLS} \xrightarrow{p} \beta$ only if $H_0$ is true but is more efficient than $\hat\beta_{IV}$.  

assuming homoskedasticity:
Then we have $\hat {var}(\hat\beta^{IV}-\hat\beta^{OLS})= var(\hat\beta^{IV})-var(\hat\beta^{OLS})$  
as $cov(\hat \beta^{IV},\hat \beta^{OLS})=var(\hat \beta^{OLS})$  

$H=(\hat \beta^{IV}-\hat \beta^{OLS})'[\hat{var}(\hat \beta^{IV})-\hat{var}(\hat \beta^{OLS})]^{-1}(\hat \beta^{IV}-\hat \beta^{OLS})$  

In a single parameter case, we have $H= \frac{(\hat \beta^{IV}-\hat \beta^{OLS})^2}{\hat{var}(\hat \beta^{IV})-\hat{var}(\hat \beta^{OLS})}$ 


##Hausman results

```{r}


# excluding lnw from OLS
#homoscedastic case
OLSS=lm(ln_qs~ln_fp+trendsp1+trendsp2+trendsp3,dta)
IVS <- ivreg(ln_qs~ln_fp+ln_w+trendsp1+trendsp2+trendsp3|ln_w+trendsp1+trendsp2+trendsp3+lag_ln_w,data=dta)

b_iv=summary(IVS)$coef[2,1]
var_b_iv=(summary(IVS)$coef[2,2])^2

b_ols=summary(OLSS)$coef[2,1]
var_b_ols=(summary(OLSS)$coef[2,2])^2

var_delta_b=var_b_iv-var_b_ols
H=((b_iv-b_ols)^2)/var_delta_b

critical_x=pchisq(0.95,1)
critical_F=pchisq(0.95,1,45)

#h2 <- hausman.systemfit( fit2sls, fit3sls ) 

# including lnw in OLS
OLSS2=lm(ln_qs~ln_fp+ln_w+trendsp1+trendsp2+trendsp3,dta)

b_ols2=summary(OLSS2)$coef[2,1]
var_b_ols2=(summary(OLSS2)$coef[2,2])^2
var_delta_b2=var_b_iv-var_b_ols2
H2=((b_iv-b_ols2)^2)/var_delta_b2
```

Our Hausman results testing the OLS without ln(w) against the IV:

$$H= -44.393$$

And the results testing the OLS with ln(w) against the IV:

$$H= 0.076$$
$$c_x = 0.670,  c_F=4.92\times10^{-9}$$ 
We reject the null hypothesis that price is exogenous in the first case and fail to reject in the second.  

##Supply Elasticity for Area

```{r}
###4 Repeat supply model with ln(area) as dependent variable
###ln_w is not included 
dta <- dta %>% mutate(ln_area=log(area)) 

olsarea=lm(ln_area~ln_fp+ln_w+trendsp1+trendsp2+trendsp3,dta) %>% tidy
IVSt <- ivreg(ln_qs~ln_fp+ln_w+trendsp1+trendsp2+trendsp3|ln_w+trendsp1+trendsp2+trendsp3+lag_ln_w,data=dta)%>%tidy
IVarea <- ivreg(ln_area~ln_fp+ln_w+trendsp1+trendsp2+trendsp3|ln_w+trendsp1+trendsp2+trendsp3+lag_ln_w,data=dta) %>% tidy

tIVS <- IVSt[-c(1, 3, 4, 5, 6), ]
tolsarea <- olsarea[-c(1, 3, 4, 5, 6), ]
tIVarea <- IVarea[-c(1, 3, 4, 5, 6, 7), ]

tablefor4 <- rbind(tOLSS2, tolsarea, tIVS, tIVarea)
rownames(tablefor4)= c("OLS of ln_qs with ln_w","OLS of ln_area with ln_w","IV of ln_qs with ln_w", "IV of ln_area with ln_w")

kable(tablefor4)

```

We can observe that the difference between the full quantity and just the area is similar in OLS and IV.


## Adding 7 Years

```{r}
###5 adding the 7 extra years doesn't seem to change the IV estimates for the supply and demand model. 
###However, it did seem to affect our Hausmen test results. Whether Ho: price is exogenous or not.
OLSSt_5=lm(ln_qs~ln_fp+trendsp1+trendsp2+trendsp3,dta2)%>% tidy
tOLSS_5 <- OLSSt_5[-c(1, 3, 4, 5), ]
OLSS2t_5=lm(ln_qs~ln_fp+ln_w+trendsp1+trendsp2+trendsp3,dta2)%>% tidy
tOLSS2_5 <- OLSS2t_5[-c(1, 3, 4, 5, 6), ]
#a=lm(ln_qs~ln_fp+ln_w+trendsp1+trendsp2+trendsp3,dta)
##demand
OLSDt_5=lm(ln_qd~ln_sp+trendsp1+trendsp2+trendsp3,dta2)%>% tidy
tOLSD_5 <- OLSDt_5[-c(1, 3, 4, 5), ]
## dropping the ln_w term gives us somewhat closer value

IVSt_5 <- ivreg(ln_qs~ln_fp+ln_w+trendsp1+trendsp2+trendsp3|ln_w+trendsp1+trendsp2+trendsp3+lag_ln_w,data=dta2)%>% tidy
tIVS_5<- IVSt_5[-c(1, 3, 4, 5, 6), ]

IVDt_5 <- ivreg(ln_qd~ln_sp+trendsp1+trendsp2+trendsp3|trendsp1+trendsp2+trendsp3+ln_w,data=dta2)%>% tidy
tIVD_5 <- IVDt_5[-c(1, 3, 4, 5, 6), ]


OLSIVtable1 <- rbind(tOLSS_5, tOLSS2_5, tOLSD_5, tIVS_5, tIVD_5)
rownames(OLSIVtable1)= c("OLS S", "OLS S with ln(w)", "OLS D", "IV S", "IV D")
table5 <- rbind(OLSIVtable,OLSIVtable1)
kable(table5)
```

Adding the 7 extra years doesn't seem to change the IV estimates for the supply and demand model. 

##Published Data

It is inconvenient to post data in .dta format, since researchers who do not use stata are then unable to replicate the work.


Otherwise, their data organization is good.  They include:  
- raw data, study datasets, and .do files to show changes  
- clear labels on data within stata  
- .do files for regressions and figures with clear comments and notes

Their work is clearly documented and could be replicated.  

