---
title: "Assignment 7"
author: "USER"
date: "March 6, 2018"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)
```

## Replication of RS' main results
```{r IV and OLS}
library(tidyverse)
# install.packages("AER")
library(AER)
library(knitr)
rm(list=ls())

# data.all = read.csv("C:/Users/tluo/Downloads/RSdata.csv")  # Location of downloaded dataset
data.all <- read_csv("https://raw.githubusercontent.com/ucdavis/ARE_ECON_240A_2018/master/Computing_Assignment_VII/data/RSdata.csv")

# Can we obtain the data using TA's method?
RSdata.ext = data.all %>% mutate(lag_ln_w=lag(ln_w), ln_area=log(area)) %>% filter(year>=1962)
RSdata = filter(RSdata.ext, year<=2007)

# OLS #
OLS11 = lm(ln_qs ~ ln_fp + ln_w + trendsp1 + trendsp2 + trendsp3, data=RSdata) #include ln_w
OLS12 = lm(ln_qs ~ ln_fp + trendsp1 + trendsp2 + trendsp3, data=RSdata) #exlude ln_w
OLS2 = lm(ln_qd ~ ln_sp + trendsp1 + trendsp2 + trendsp3, data=RSdata)
# IV #
iv1 = ivreg(ln_qs ~ ln_fp + ln_w + trendsp1 + trendsp2 + trendsp3 |
                    ln_w + lag_ln_w + trendsp1 + trendsp2 + trendsp3, data=RSdata)
                    # include ln_w in both equations
iv2 = ivreg(ln_qd ~ ln_sp + trendsp1 + trendsp2 + trendsp3 |
                    ln_w + trendsp1 + trendsp2 + trendsp3, data=RSdata)

# OLS results #
kable(rbind(ln_fp.include = summary(OLS11)$coef["ln_fp",-3], ln_fp.exclude =    summary(OLS12)$coef["ln_fp",-3], ln_sp = summary(OLS2)$coef["ln_sp",-3]), digits = 3, caption = "Table 1a. OLS Estimates")  
# IV results #
kable(rbind(ln_fp = summary(iv1)$coef["ln_fp",-3], ln_sp = summary(iv2)$coef["ln_sp",-3]), digits = 3, caption="Table 1b. IV Estimates")

```
  
(1) RS excluded ln_w in their OLS model for the supply side.  
(2) The estimates do not match exactly to the results of the paper since our data of ln_w is a little bit different from RS's but still compatible with RS's results in terms of signs, magnititudes and significances.  

## Hausman Test
Suppose that  
$$y_i = x_i'\beta + e_i, \quad E[x_i e_i]=0 (H_0); \quad x_i = \Gamma' z_i + v_i, \quad E[z_i v_i] = E[z_i e_i] = 0$$
Then
$$E[x_i e_i] = E[(\Gamma' z_i + v_i) e_i] = \Gamma'E[z_i e_i] + E[v_i e_i] =0$$
If $E[x_i e_i] \neq 0$, then $E[v_i e_i] \neq 0$. Thus, a linear regression model with OLS can be used to test whether the projection coefficient of $e_i$ over $v_i$ is zero or not as a method to test $E[x_i e_i]=0$. 

```{r Hausman}
## Hausman Test - Existing test##
hausman = summary(iv1, diagnostics = T)$diagnostics["Wu-Hausman","p-value"]

## Hausman Test- Manual ##
#step 1: obtain the residuals of the first stage of IV regression
resid = lm(ln_fp ~ ln_w + lag_ln_w + trendsp1 + trendsp2 + trendsp3, data=RSdata)$res

# step 2: 
ols1 = lm(ln_qs ~ ln_fp + trendsp1 + trendsp2 + trendsp3 + resid, 
          data=cbind(RSdata,resid)) # exclude ln_w
ols2 = lm(ln_qs ~ ln_fp + ln_w + trendsp1 + trendsp2 + trendsp3 + resid,    
          data=cbind(RSdata,resid)) # include ln_w
##manual vs. existing
kable(data.frame(ols.exlude = summary(ols1)$coef["resid","Pr(>|t|)"], ols.include = 
      summary(ols2)$coef["resid","Pr(>|t|)"], hausman), digits=3, caption= "Husman Test")
```

(1) If ln_w included, there is no significant difference between OLS and IV.  
(2) If ln_w excluded, there is significant difference between OLS and IV.  

## Supplemental results

1. Supply elasticity for area    

```{r Area}
iv3 = ivreg(ln_area ~ ln_fp + ln_w + trendsp1 + trendsp2 + trendsp3 |
            ln_w + lag_ln_w + trendsp1 + trendsp2 + trendsp3, data=RSdata)

kable(t(data.frame(ln_fp=summary(iv3)$coef["ln_fp",-3])), digits=3, caption = "Table 3. Supply elasticity for Area")
```

The supply elasticity for area is lower than that for total quantity but not that much. In fact, they are comparable in terms of sign, magnititude and significance.  

2. Impact of additional years of data  

```{r ext}
iv1b = ivreg(ln_qs ~ ln_fp + ln_w + trendsp1 + trendsp2 + trendsp3 | ln_w +     
             lag_ln_w + trendsp1 + trendsp2 + trendsp3, data=RSdata.ext)
iv2b = ivreg(ln_qd ~ ln_sp + trendsp1 + trendsp2 + trendsp3 | ln_w + 
             trendsp1 + trendsp2 + trendsp3, data=RSdata.ext)

kable(rbind(ln_fp = summary(iv1b)$coef["ln_fp",-3], ln_sp = summary(iv2b)$coef["ln_sp",-3]), digits = 3, caption = "Table 4. Additional 7 years of recent data")

```

Estimates have the same signs as before with only slight differences in terms of magnitudes and significances.  

## Comments on original datasets
The orginal datasets and code are provided by the authors in a modular fashion. The authors could have organized the data in a single dataset and provided the code in only one file with indicated sections, which would have made replicating the results a bit easier. But on the other hand, compared to many other authors who provide data and code with little explaination, theirs is relatively complete. 