---
title: "Computing Assignment VII"
author: "group"
date: "March 9, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(AER)

RSdata_raw <- read_csv("https://raw.githubusercontent.com/ucdavis/ARE_ECON_240A_2018/master/Computing_Assignment_VII/data/RSdata.csv")%>% mutate(.,ln_area=log(area))
RSdata <- na.omit(RSdata_raw)
##1
reg_qs<- lm(ln_qs ~ ln_fp+trendsp1+trendsp2+trendsp3, data=RSdata,na.rm=TRUE)


reg_qd<- lm(ln_qd ~ ln_sp +trendsp1+ trendsp2+trendsp3, data=RSdata)
summary(reg_qs)
summary(reg_qd)

##2
IV_qs<-ivreg(ln_qs~ ln_fp + ln_w+ trendsp1+trendsp2 +trendsp3 | ln_w+lag_ln_w+ trendsp1+trendsp2+trendsp3, data=RSdata)
IV_qd<-ivreg(ln_qd~ ln_sp +trendsp1+ trendsp2+trendsp3 |ln_w+ trendsp1+trendsp2+trendsp3, data=RSdata)

summary(IV_qs)
summary(IV_qd)


##4

IV_area<-ivreg(ln_area~ ln_fp + ln_w+ trendsp1+trendsp2 +trendsp3 | ln_fp+ln_w+lag_ln_w+ trendsp1+trendsp2+trendsp3, data=RSdata)
summary(IV_area)
summary(IV_qs)

```

The supply elasticity for total quantity is approximately $0.9$ and that for area is $0.63$. The difference stems from $log(yield)$ which is $ln(qs) - ln(area)$. The mean of $log(yield)$ is $2.13$ ranging from $1.57$ to $2.57$. The correlation of $ln(qs)$ and $log(area)$ is almost one.

```{r setup, include=FALSE}

cor(RSdata$ln_area, RSdata$ln_qs)
a<-RSdata$ln_qs-RSdata$ln_area
max(a)
mean(a)
min(RSdata$ln_qs)
min(a)

##5

RSdata_reduced<- filter(RSdata, year<=2008)
IV_qs_reduced<-ivreg(ln_qs~ ln_fp + ln_w+ trendsp1+trendsp2 +trendsp3 | ln_w+lag_ln_w+ trendsp1+trendsp2+trendsp3, data=RSdata_reduced)
IV_qd_reduced<-ivreg(ln_qd~ ln_sp +trendsp1+ trendsp2+trendsp3 |ln_w+ trendsp1+trendsp2+trendsp3, data=RSdata_reduced)

summary(IV_qs_reduced)  
```

The supply elasticity is almost $0.1$, which is close to the coefficient estimated with the full sample.

```{r setup, include=FALSE}
summary(IV_qd_reduced)  

```

The demand elasticity is almost $-0.066$, which is close to the coefficient $(-0.063)$ estimated with the full sample.


