---
title: "Computing Assignment VII"
author: "group"
date: "March 9, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(AER)

# Problem 1
RSdata_raw <- read_csv("https://raw.githubusercontent.com/ucdavis/ARE_ECON_240A_2018/master/Computing_Assignment_VII/data/RSdata.csv")%>% mutate(.,ln_area=log(area))
RSdata <- na.omit(RSdata_raw)

reg_qs<- lm(ln_qs ~ ln_fp+trendsp1+trendsp2+trendsp3, data=RSdata,na.rm=TRUE)
reg_qd<- lm(ln_qd ~ ln_sp +trendsp1+ trendsp2+trendsp3, data=RSdata)
summary(reg_qs)
summary(reg_qd)

# Problem 2
IV_qs<-ivreg(ln_qs~ ln_fp + ln_w+ trendsp1+trendsp2 +trendsp3 | ln_w+lag_ln_w+ trendsp1+trendsp2+trendsp3, data=RSdata)
IV_qd<-ivreg(ln_qd~ ln_sp +trendsp1+ trendsp2+trendsp3 |ln_w+ trendsp1+trendsp2+trendsp3, data=RSdata)

summary(IV_qs)
summary(IV_qd)

# Problem 3

# Excluding "ln_w."
reg_3_ex_1st <- lm(ln_fp ~ lag_ln_w + ln_w + trendsp1 + trendsp2 + trendsp3, data = RSdata)
reg_3_ex_residual <- residuals(reg_3_ex_1st) %>% as.data.frame() %>% `colnames<-`(c("reg_3_ex_residual"))
RSdata <- cbind(RSdata, reg_3_ex_residual)

reg_3_ex_2nd <- lm(ln_qs ~ ln_fp + reg_3_ex_residual + trendsp1 + trendsp2 + trendsp3, data = RSdata)

# Including "ln_w."
reg_3_in_1st <- lm(ln_fp ~ lag_ln_w + ln_w + trendsp1 + trendsp2 + trendsp3, data = RSdata)
reg_3_in_residual <- residuals(reg_3_in_1st) %>% as.data.frame() %>% `colnames<-`(c("reg_3_in_residual"))
RSdata <- cbind(RSdata, reg_3_in_residual)

reg_3_in_2nd <- lm(ln_qs ~ ln_fp + ln_w + reg_3_in_residual + trendsp1 + trendsp2 + trendsp3, data = RSdata)


summary(reg_3_ex_2nd)
summary(reg_3_in_2nd)
```
We can now calculate a Hausman test statistic for the cases of including and excluding ln_w from the OLS regression compared to the IV regression, to see whether endogeneity is present in either case. Let $\hat{\beta}_1$ represent OLS coefficient without ln_w and $\hat{\beta}_2$ represent OLS coefficient with ln_2. Let $\tilde{\beta}$ represent IV coefficient. 

Our first test is:
$H_O$: $H = ($\hat{\beta}_1$- $\tilde{\beta}$)^2/(\hat{s}_1^2 - \tilde{s}^2) = 0$
$H_a$: $H \neq 0$

Using results from our OLS and IV regression we get a test statistic of $H = 194.153$, which is $\chi^2(1)$ distributed, and is much higher than the critical value of 3.84. Thus we reject the null hypothesis and suspect that endogeneity is present.

Our second test is:
$H_O$: $H = ($\hat{\beta}_2$- $\tilde{\beta}$)^2/(\hat{s}_2^2 - \tilde{s}^2) = 0$
$H_a$: $H \neq 0$

The result is H = 0.48, which is less than critical value 3.84, thus we do not reject the null hypothesis, and we have less evidence of endogeneity.

We can also carry out the Hausman test another way (as seen in our regression approach) by regressing supply on the residuals of the original regression (either with or without ln_w) and testing whether the coefficient on the residuals is different from zero. Through this approach we find equivalent results, with $t = 2.99 > 1.96$ when ln_w is excluded and $t = -0.71 > -1.96$ when ln_w is included in the OLS regression. 

Thus we concluded that controlling for supply shock in the OLS regression reduces the endogeneity problem of measuring the effect of futures price on output supply.

```{r setup2, include=FALSE}
# Problem 4

IV_area<-ivreg(ln_area~ ln_fp + ln_w+ trendsp1+trendsp2 +trendsp3 | ln_fp+ln_w+lag_ln_w+ trendsp1+trendsp2+trendsp3, data=RSdata)
summary(IV_area)
summary(IV_qs)

```

The supply elasticity for total quantity is approximately $0.9$ and that for area is $0.63$. The difference stems from $log(yield)$ which is $ln(qs) - ln(area)$. The mean of $log(yield)$ is $2.13$ ranging from $1.57$ to $2.57$. The correlation of $ln(qs)$ and $log(area)$ is almost one.

```{r, include=FALSE}

cor(RSdata$ln_area, RSdata$ln_qs)
a<-RSdata$ln_qs-RSdata$ln_area
max(a)
mean(a)
min(RSdata$ln_qs)
min(a)

# Problem 5

RSdata_reduced<- filter(RSdata, year<=2008)
IV_qs_reduced<-ivreg(ln_qs~ ln_fp + ln_w+ trendsp1+trendsp2 +trendsp3 | ln_w+lag_ln_w+ trendsp1+trendsp2+trendsp3, data=RSdata_reduced)
IV_qd_reduced<-ivreg(ln_qd~ ln_sp +trendsp1+ trendsp2+trendsp3 |ln_w+ trendsp1+trendsp2+trendsp3, data=RSdata_reduced)

summary(IV_qs_reduced)  
```

The supply elasticity is almost $0.1$, which is close to the coefficient estimated with the full sample.

```{r, include=FALSE}
summary(IV_qd_reduced)  

```

The demand elasticity is almost $-0.066$, which is close to the coefficient $(-0.063)$ estimated with the full sample.


