---
title: "CAIII"
author: "OMG"
date: "February 1, 2018"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Exercise 1: Leverage Plot

```{r cars, echo=FALSE, message=FALSE, warning=FALSE}
library(dplyr)
library(sandwich)
library(lmtest)
library(tidyverse)
library(ggpubr)
#Exercise 1
simulator<-function(n,distrib){
  set.seed(12345)
  te=rnorm(n)
  if (identical(distrib,"u")) {tx=runif(n) 
  distr="uniform"}
  if (identical(distrib,"n")) {tx=rnorm(n)
  distr="normal"}
  if (identical(distrib,"q")) {tx=rchisq(n,1) 
  distr="chisq"}
  
  ty=(tx^2)+te
  tsim=data.frame(ty,tx)
  lmodel <- lm(ty~ tx, data = tsim)
  leverage <- hatvalues(lmodel)
  tsim <- data.frame(ty,tx,leverage)
  summary(lmodel)
  
  sim_n1=c(coef(summary(lmodel))[1, 1], coef(summary(lmodel))[2, 1])
  return(tsim)

}
#run the simulation of sample size of 100 once  
tsim <- simulator(100,"q")
#graph the scatter plot
G1 <- ggplot(aes(x=tx, y=ty, size=leverage), data=tsim) + geom_point()
#Leave out the observations with large leverage
x_small <- tsim[(tsim[,"leverage"]<0.1),] 
#graph again
G2 <- ggplot(aes(x=tx, y=ty, size=leverage), data=x_small) + geom_point()
plot(G1)
```

## Exercise 2

**Replicated table**

Comments on results of the oracle estimate:

- We can see that for the heteroscedastic case the rejection rate with the oracle computation of the var-cov matrix is closer to 5%.
- We use the same t distribution with 28 degrees of freedom, hence the error in the rejection rate of "constant" and "HC1" comes from the estimation of the var-cov matrix.
- For the homoscedastic case, the best estimator is the "constant" methodology (asumming homoskedasticity)
- However, doing oracle doesnt significantly affect the rejection rate.

```{r pressure, echo=FALSE, message=FALSE, warning=FALSE}
library(ggpubr)

simulate<-function(sig){
#sample size
n=30
#generate error
h_var=sig
D=rep(0,n)
D[1:3]=1
b0=0
b1=0
e=rep(1,n)
e[1:3]=rnorm(3)
e[4:30]=sqrt(h_var)*rnorm(27)
#or we could just do rnorm(27,0,h_var)?
y=b0+b1*D+e
tsim=data.frame(y, D)

lmodel <- lm(y~ D, data = tsim)
#homoscdedastic #I thought homo should be constant no?
#I thought the default they are using in R is HC3
#ct1 <- coeftest(lmodel,  vcov = vcovHC(lmodel,"const"))
output1=coef(summary(lmodel))[2,]
#add robust estimate of se
output2=coeftest(lmodel, df = 28, vcov = vcovHC(lmodel, type = "HC1"))[2,]
#input the true v-cov matrix of errors
sigmas=c(rep(1,3), rep(sig,27))
output3=coeftest(lmodel, df = 28, vcov = sandwich(lmodel,meat.=meatHC(lmodel,omega=sigmas)))[2,]
output=data.frame(output1,output2,output3)
reject=output[4,]<0.05 #I think we also need a rejection rate somewhere,I added something here
output=rbind(output,reject)
names(output)<-c("const", "HC1","unf")
rownames(output)[5]<-"reject"
return(output)

}


S <- 10000
params_gr <- expand.grid(sim=1:S, sig=c(0.5, 1))
expaned=data.frame(params_gr,freq=5)
t=params_gr[rep(seq_len(nrow(params_gr)), each=5),]
res_gr <- mapply(simulate, sig=params_gr$sig, SIMPLIFY = FALSE)%>%
  bind_rows()
var=rep(c("est","se", "t", "pi","reject"), 20000)
complete=cbind(t,var,res_gr)


#subetting example
#sig == 0.5
mc1 <- mean(complete$const[which(complete$var=="se" & complete$sig==0.5)])
sc1 <- sd(complete$const[which(complete$var=="se" & complete$sig==0.5)])
mh1 <- mean(complete$HC1[which(complete$var=="se" & complete$sig==0.5)])
sh1 <- sd(complete$HC1[which(complete$var=="se" & complete$sig==0.5)])
rc1 <- mean(complete$const[which(complete$var=="reject" & complete$sig==0.5)])
rh1 <- mean(complete$HC1[which(complete$var=="reject" & complete$sig==0.5)])
#sig == 1
mc2 <- mean(complete$const[which(complete$var=="se" & complete$sig==1)])
sc2 <- sd(complete$const[which(complete$var=="se" & complete$sig==1)])
mh2 <- mean(complete$HC1[which(complete$var=="se" & complete$sig==1)])
sh2 <- sd(complete$HC1[which(complete$var=="se" & complete$sig==1)])
rc2 <- mean(complete$const[which(complete$var=="reject" & complete$sig==1)])
rh2 <- mean(complete$HC1[which(complete$var=="reject" & complete$sig==1)])

#sig==0.5 unf
mc3 <- mean(complete$unf[which(complete$var=="se" & complete$sig==0.5)])
sc3 <- sd(complete$unf[which(complete$var=="se" & complete$sig==0.5)])
mh3 <- mean(complete$unf[which(complete$var=="se" & complete$sig==1)])
sh3 <- sd(complete$unf[which(complete$var=="se" & complete$sig==1)])
rc3 <- mean(complete$unf[which(complete$var=="reject" & complete$sig==0.5)])
rh3 <- mean(complete$unf[which(complete$var=="reject" & complete$sig==1)])

l1 <- c(mc1,sc1,rc1)
l2 <- c(mh1,sh1,rh1)
l3 <- c(mc2,sc2,rc2)
l4 <- c(mh2,sh2,rh2)
l5 <- c(mc3,sc3,rc3)
l6 <- c(mh3,sh3,rh3)
table<- rbind(l1,l2,l3,l4,l5,l6)

table2<-rbind(l1,l2,l3,l4)

rownames(table)=c("Ht err- hom est", "Ht err- HC1", "Hom err- hom est", "Hom err- HC1","Ht err-oracle","Hom err-oracle") 
colnames(table)=c("mean", "sd", "r rate")

homoconst <- complete$const[which(complete$var=="t", complete$sig==1)]
homoHC <- complete$HC1[which(complete$var=="t", complete$sig==1)]
heteroconst <- complete$const[which(complete$var=="t", complete$sig==0.5)]
heteroHC <- complete$HC1[which(complete$var=="t", complete$sig==0.5)]
graph <- data.frame(homoconst,homoHC,heteroconst,heteroHC)
A <- ggplot(aes(x=homoconst), data=graph) +geom_density()+ stat_function(fun = dt, args = list(df = 28), colour = "red")
B <- ggplot(aes(x=homoHC), data=graph) +geom_density()+ stat_function(fun = dt, args = list(df = 28), colour = "red")
C <- ggplot(aes(x=heteroconst), data=graph) +geom_density()+ stat_function(fun = dt, args = list(df = 28), colour = "red")
D <- ggplot(aes(x=heteroHC), data=graph) +geom_density()+ stat_function(fun = dt, args = list(df = 28), colour = "red")
Final <- ggarrange(A,B,C,D + rremove("x.text"), labels = c("A", "B", "C", "D"),ncol = 2, nrow = 2)

print(table)
```
## Combined plots for the t-ratio
```{r, echo=FALSE}
plot(Final)
```

