---
title: "Assignment III"
author: "YEP"
date: "2/1/2018"
output: beamer_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)
```

## Leverage plot
As expected, observations with larger x have higher leverage:

```{r }
library(ggplot2)
library(dplyr)
library(sandwich)

set.seed(102030)

  x <-rchisq(100,df=1)
  e <-rnorm(100)
  y <- x^2+e
  b <- lm(y~x, data=as.data.frame(cbind(x,e,y)))

leverage <- hatvalues(b)
data <- data.frame(cbind(x,y,leverage))

 ggplot(aes(x=x, y=y, size=leverage),data=data) +
   geom_point()
```

## OLS estimator with HC covariance
```{r }
dgp <- function(sig){
   d <-c(rep(1,3),rep(0,27))
  
   sigmas <-c(rep(1, 3),rep(sig, 27))
   e <-rnorm(30,mean = 0,sd = sigmas)
   y <- e
   reg <- lm(y~d, data=as.data.frame(cbind(y,d)))
   conv <- cbind(sig,"conventional",t(summary(reg)$coef[2,]))
   
   HC1 <- vcovHC(reg, type = "HC1")[2,2]
   HC1 <- cbind(sig,"HC1",conv[3],sqrt(HC1),0,0)
   
   uHC1 <- vcovHC(reg, omega= (30/28)*sigmas^2)[2,2]
   uHC1 <- cbind(sig,"uHC1",conv[3],sqrt(uHC1),0,0)
   
   res <- data.frame(rbind(conv,HC1,uHC1))
   
   return(res)
 }
 

 S <- 10000
 params_gr <-expand.grid(1:S, sig=c(0.5, 1))
 res_gr <-mapply(dgp, sig=params_gr$sig, SIMPLIFY = FALSE) %>%
   bind_rows()
 res_gr$Std..Error <- as.numeric(res_gr$Std..Error)
 res_gr$Estimate <- as.numeric(res_gr$Estimate)
 res_gr$Pr...t.. <- as.numeric(res_gr$Pr...t..)
 res_gr$tt<- res_gr$Estimate/res_gr$Std..Error 
 res_gr$p_values <- pt(-abs(res_gr$tt),28)*2
 
 
 res_gr %>% group_by(sig) %>% filter(V2== "conventional") %>% 
   summarise(mean=mean(Estimate),sd=sd(Estimate),five=sum(Pr...t..<0.05)/S)
 
 
 res_gr %>% group_by(sig,V2) %>% 
   summarise(mean=mean(Std..Error,na.rm=T),sd=sd(Std..Error),five=sum(p_values<0.05)/S)
 
 
```

## OLS estimator with HC covariance
With heteroskedasticity, rejection rates are much larger than 5%. The estimation of standard errors are below the unfeasible HC1, but HC1 is better than the conventional.
Under homoskedasticity, the conventional behaves well, as expected.

## OLS estimator with HC covariance
```{r }
tdist.conventional.ho <- res_gr %>% filter(V2== "conventional", sig==1) %>% select(tt)
 tdist.conventional.he <- res_gr %>% filter(V2== "conventional", sig==0.5) %>% select(tt)
 tdist.HC1.ho <- res_gr %>% filter(V2== "HC1", sig==1) %>% select(tt)
 tdist.HC1.he <- res_gr %>% filter(V2== "HC1", sig==0.5) %>% select(tt)
 
 
 ggplot(data.frame(x = c(-4, 4)), aes(x = x)) +
   stat_function(fun = dt, args = list(df = 28),aes(colour = "True")) + 
   geom_density(data=tdist.conventional.ho, aes(x=tt,colour = "Conventional")) + 
   geom_density(data=tdist.HC1.ho, aes(x=tt,colour = "HC1")) + 
   theme_bw() +
   ggtitle("homoskedastic")
```
 
## OLS estimator with HC covariance
 
```{r }
 ggplot(data.frame(x = c(-4, 4)), aes(x = x)) +
   stat_function(fun = dt, args = list(df = 28),aes(colour = "True")) + 
   geom_density(data=tdist.conventional.he, aes(x=tt,colour = "Conventional")) + 
   geom_density(data=tdist.HC1.he, aes(x=tt,colour = "HC1")) + 
   theme_bw() +
   ggtitle("heteroskedastic")

```