---
title: "Assign7"
author: "6PM"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r , results = "hide", message=FALSE}
library(tidyverse) #everything
library(lmtest) # coeftest
library(sandwich) # vcovHC
library(dplyr)
library(broom) # tidy the output
library(knitr) # For knitting document and include_graphics function
library(AER) # ivreg
library(plyr)
library(memisc)
rm(list = ls())      # clean the workspace
data <- read.csv(file = "/Volumes/Hugo/Study/UCD/Classes/ECN240A/PS7/RSdata.csv", header = TRUE, stringsAsFactors = F, na.strings = ".")
str(data)
s_ols <- lm(ln_qs ~ ln_fp + ln_w + trendsp1 + trendsp2 + trendsp3, 
            data = data)
d_ols <- lm(ln_qd ~ ln_sp + trendsp1 + trendsp2 + trendsp3, 
            data = data)
s_ols_nw <- lm(ln_qs ~ ln_fp + trendsp1 + trendsp2 + trendsp3, data = data)
# IV regression
data<-data%>%
  mutate(lag_ln_w = dplyr::lag(ln_w))
s_iv <- ivreg(ln_qs ~ ln_fp + ln_w + trendsp1 + trendsp2 + trendsp3 | .-ln_fp + lag_ln_w, data = data)
d_iv <- ivreg(ln_qd ~ ln_sp + trendsp1 + trendsp2 + trendsp3 | .-ln_sp + ln_w, data = data)
s_iv_nw <- ivreg(ln_qs ~ ln_fp + trendsp1 + trendsp2 + trendsp3 | .-ln_fp + lag_ln_w, data = data)
first_stage <- lm(ln_fp ~ lag_ln_w + ln_w + trendsp1 + trendsp2 + trendsp3, data = data)

# using area instead
data<-data%>%
  mutate(ln_area = log(area))
s_iv_area <- ivreg(ln_area ~ ln_fp + ln_w + trendsp1 + trendsp2 + trendsp3 | .-ln_fp + lag_ln_w, data = data)
```

## 1. OLS Regression
```{r ols}
mtable('Supply 1' = s_ols_nw,
       'Supply 2' = s_ols,
       'Demand' = d_ols,
       summary.stats=c("adj. R-squared", "F", "p", "N"),show.eqnames=FALSE)
```

## Comparison with the paper
- For supply elasticity, our OLS estimate of $\beta_s$ with ln_w is 0.089(0.012), which is statistically significant. 
- However, without ln_w it's 0.038(0.024), close to what RS did (0.02(0.031)), both are not statistically significant. Due to the large discrepancy, we guess RS used the one without ln_w. 
- For demand elasticity, our OLS estimate of $\beta_d$ is -0.017(0.011), close to what RS got -.018(0.010), but ours is not significant. 

## 2. IV Regression
```{r iv}
mtable('First Stage' = first_stage,
       'Supply' = s_iv,
       'Demand' = d_iv,
       summary.stats=c("adj. R-squared", "F", "Wald", "p", "N"),
       show.eqnames=FALSE)
```

## Comparison with the paper
- For supply elasticity, our IV estimate of $\beta_s$ with $ln_w$ is 0.102(0.025), which is statistically significant, close to what RS got 0.096(0.025).
- For demand elasticity, our IV estimate of $\beta_d$ is -0.063(0.0228), close to what RS got -0.055(0.024), and both are significant at 5% confidence level. 

## 3. Hausman Test
```{r hausman, results = "hide", message=FALSE}
test <- function(a,b){
  beta_a = summary(a)$coefficients[2,1]
  se_a = summary(a)$coefficients[2,2]
  beta_b = summary(b)$coefficients[2,1]
  se_b = summary(b)$coefficients[2,2]
  cf_diff <- beta_a - beta_b
  vc_diff <- se_a^2 - se_b^2
  Hausman <- cf_diff^2/vc_diff
  p <- pchisq(Hausman, df = 1, lower.tail = FALSE)
  test <- data_frame(
    Hausman = Hausman,
    p = round(p, digits = 3))
  return(test)
}
```

Comparison between IV and OLS with ln_w
```{r test1}
test(s_iv,s_ols)
```

Comparison between IV and OLS without ln_w
```{r test2}
test(s_iv,s_ols_nw)
# Compared with Wu-Hausman test
# summary(s_iv, diagnostics = TRUE)
# summary(s_iv_nw, diagnostics = TRUE)
```

## Comments on the result of Hausman test
- We find that with ln_w in the OLS regression, the p value of Hausman test is large enough so that we can't reject the null, there's no big difference between OLS and IV estimates. 
- On the other hand, without ln_w in the OLS regression, the p value is almost zero, we can reject the null, and IV estimate is pretty different from OLS estimate. 
- Based on these results, our explanation is as follows: 
From the first stage regression, we know that there's a strong negative relationship between ln_fp and ln_w, so if we omit ln_w in the OLS regression, the residual should be correlated to ln_fp, which will generate endogeneity bias. In this case, since ln_w also positively correlated with ln_qs, $\beta_s$ will bias downward (using omitted variable bias formula).
- After controlling ln_w, according to the Hausman test, there's no big difference between OLS and IV estimates. One possible reason is that there's little correlation between error term and ln_fp, thus no need to apply IV regression; another possible reason is that maybe lag_ln_w is not a good IV, there's something left in the error term still correlates with our IV. 

## Second thoughts
- We did something more, we also look at the Hausman test comparing OLS and IV both without ln_w. We find no significant difference between them. If lag_ln_w is a good IV, since ln_fp and ln_w are highly correlated, there exists endogeneity problem, then doing IV regression should help (given that lag_ln_w is not highly correlated with ln_w), we should expect some difference there, but No! Thus this result implies that lag_ln_w seems to be a bad IV. 

Comparison 
```{r nw2}
test(s_iv_nw,s_ols_nw)
summary(lm(ln_w ~ lag_ln_w - 1,data = data))
```

## 4. Is the supply elasticity different for area and total quantity?
```{r iv_area}
summary(s_iv_area)
```

- What we find is that the IV estimate of supply elasticity $\beta_s$ for area is 0.077(0.021), which is statistically significant. Comparing with the IV estimate of supply elasticity for total quantity 0.096(0.025), we can see that most of the price effect on the supply side lies in the change of grow areas (extensive margin). 

## 5. Extra data helps?
- It seems that the regression results are pretty close to what RS got in their paper.

## 6. Comments on the original data
- There are multiple folders containing all their work including data process and estimations. Since the STATA can only process one dataset at once, combining data from different datasets takes a lot of time.
We spent 20 minutes doing the createData.do, and nothing done. So we quit. Besides, it's difficult to understand their structure without reading their instruction. But we admire their efforts!