---
title: "Assignment 4"
author: "USER"
date: "February 4, 2018"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)
```

## Convergence of the OLS estimator

Two types of convergence for the OLS estimator as the sample size goes to infinity:   
 1. Point: the estimator, Beta hat, converges to a constant, the BLP coefficient     
 2. Distribution: the scaled estimator, sqrt(n)*(Beta hat - Beta), converges to a normal distribution  

```{r OLS Convergence, fig.height = 4, fig.width = 10}
rm(list=ls())
set.seed(1000)

## Convergence to constant ##
N=1.2^60
ns = floor(1.2^(10:60))
x = runif(N)
y = x + rnorm(N)
dat1 = data.frame(x,y)
sim1 = cbind(n = ns, betas = sapply(ns, function(n) as.vector(lm(y ~ x, data=dat1[1:n,])$coef["x"])))

par(mfrow=c(1,2))
plot(sim1, type="l", log="x", ylab="OLS estimate", main="Convergence to Constant",col="blue")
abline(h=1, lty=2,col="red")

## Convergence to distribution ##
ns = floor(1.2^(10:40))
get.betas.1 = function(){
  N=1.2^40
  x = runif(N)
  y = x + rnorm(N)
  dat2=data.frame(x,y)
  sapply(ns, function(n) as.vector(lm(y ~ x, data=dat2[1:n,])$coef["x"]))
}
results.1 = replicate(1000, get.betas.1())
results.dist = (results.1-1)*sqrt(ns)

plot(0, type="n", xlab="sqrt(n)*(Beta hat - Beta)", ylab="Density", main="Convergence in Distribution",
     xlim=c(-10,10), ylim=c(-0,.12))
for (d in 1:length(ns)) lines(density(results.dist[d,]), col=rgb(d/length(ns),0,1-d/length(ns),.2), lwd=3)
curve(dnorm(x, sd=sqrt(12)), add=T, lwd=3, lty=2)
legend("topleft", legend=c("asymptotic distribution","n~10,","n~100,","n~1000,"),
       col=c("black",rgb(0,0,1,.2),rgb(.5,0,.5,.2),rgb(1,0,0,.2)), lty=c(2,1,1,1), lwd=rep(3,4),
       bty = "n", cex=.7)
```


## Different types of errors: Normal vs. Cauchy

When the error follows normal distribution, we have  
 1. OLS slope estimator, approaching the correponding BLP coefficient  
 2. Mean, close to the correponding BLP coefficient; Sd, close to zero  
However, when the error follows Cauchy distribution, we have  
 1. OLS slope estimator, not approaching the correponding BLP coefficient  
 2. Mean, far from the correponding BLP coefficient; Sd, far from zero  

Summary of means and SDs by n:

```{r Normal vs Cauchy- Table}
library(knitr)
rm(list=ls())
set.seed(1234)

get.betas = function(){
  N=2^13
  ns = 2^(3:13)
  x = rnorm(N)
  err = cbind(normal=rnorm(N), cauchy=rcauchy(N))
  y = 0.8 + 0.3*x + err
  dat = data.frame(y, x)
  cbind(n = ns,
        normal = sapply(ns, function(n) as.vector(lm(normal ~ x, data = dat[1:n,])$coef["x"])),
        cauchy = sapply(ns, function(n) as.vector(lm(cauchy ~ x, data = dat[1:n,])$coef["x"])))
}

results = replicate(1000, get.betas())

types = c("normal", "cauchy")
results.tab = cbind(apply(results, c(1,2), mean), apply(results, c(1,2), sd)[,-1])
colnames(results.tab) = c("n", paste(rep(types,2),rep(c("mean","sd"),each=2),sep="."))

kable(results.tab, digits = 3, caption = "Table 1: OLS Estimates")
```

## Converence for Normal vs. Cauchy (cont.)
Convergence is clear under normal, whereas cauchy does not converge even for large n
```{r Normal vs Cauchy- Plot, fig.height = 4, fig.width = 10}
### Plot ###
color.names = c("blue","red")
color.rbg = c(rgb(0,0,1,0.01),rgb(1,0,0,0.01))

par(mfrow=c(1,2))
for (t in 1:length(types)){
  bands = t(sapply(1:dim(results)[1], function(x) quantile(results[x,types[t],], c(.05,.95))))
  means = apply(results[,types[t],], 1, mean)
  plot(0, type="n", xlab="Sample size", ylab="Beta hat", main=types[t],
       ylim=range(bands)*1.5, xlim=c(8,max(results[,"n",])), log="x")
  lines(results[,c("n",types[t]),1], type="l") # One simulation set of N
  for(i in 2:dim(results)[3]) lines(results[,c("n",types[t]),i], type="l", col=color.rbg[t]) # Each sim
  for(b in 1:2) lines(results[,"n",1], bands[,b], lty=2, lwd=3, col=color.names[t]) # Interval
  lines(results[,"n",1], means, lty=3, lwd=3, col=color.names[t]) # Means
  lines(cbind(results[,"n",1],.3), lty=3, lwd=2) # True Beta
  legend("topright", legend=c("90% interval","Mean","True Beta","Example Beta path"), col=c(rep(color.names[t],2),rep("black",2)), lty=c(2,3,3,1), lwd=c(3,3,2,1), bty = "n")
}
```

Summary of means and SDs for 1 beta path:
```{r Normal vs Cauchy- Table 2}
kable(cbind(stat=c("mean","SD"),round(apply(results[,2:3,1], 2, function(x) c(mean(x),sd(x))),3)), caption="Table 2: OLS Estimates")
```
